%% chapter9.tex – Relevance
%% Group No. 2 | Dept. of CSE (Data Science), MACE Kothamangalam
%% Team: Muhammed Ajmal P (MAC23CD040), Muhammed Adil PP (MAC23CD039),
%%       Mohammed Shadeed P (MAC23CD037), Sinan Rahman (LMAC23CD066)
%% Guide: Prof. Eldo P Elias | May 2026
\chapter{Relevance}
\label{chap:relevance}

This chapter discusses the broader societal, academic, and developmental significance
of WaveGuard~V4, including its mapping to the United Nations Sustainable Development
Goals (SDGs).

% ─────────────────────────────────────────────────────────────────────────────
\section{Improving Coastal Safety}

Kerala's 590~km coastline is home to over 1.6~million fisherfolk and countless coastal
residents who face recurring threats from \textit{Kallakkadal} swell surges.
WaveGuard~V4 provides a deployable, low-cost early-warning system that fills the gap
between national-scale satellite forecasting and community-level actionable warning.
By detecting anomalous wave conditions directly in the near-shore zone and
broadcasting alerts within 8--12~seconds, the system provides residents and fishermen
sufficient time to take protective action---moving boats to harbour, securing coastal
infrastructure, and evacuating low-lying areas.

% ─────────────────────────────────────────────────────────────────────────────
\section{Helping Fishermen Make Informed Decisions}

Fishing communities in coastal Kerala frequently venture into the sea without access
to granular, near-real-time wave data.
Existing forecast services require internet connectivity and meteorological literacy
that may not be universally accessible in these communities.
WaveGuard~V4 delivers its alerts via GSM SMS---a channel requiring no smartphone or
internet access---making it immediately usable by the most at-risk segment of the
population.
The three-tier status code (CALM / MODERATE / HIGH WAVE) is deliberately simple to
interpret, enabling instant decision-making without specialist training.
A future mobile application (Section~\ref{chap:future}) will further extend
information access to the growing proportion of fishermen with smartphones.

% ─────────────────────────────────────────────────────────────────────────────
\section{Reducing Loss of Life and Property}

Each \textit{Kallakkadal} event historically results in capsized boats, destroyed
fishing equipment, property damage, and fatalities.
A five-buoy deployment at 1~km spacing could provide 5--10~minutes of advance warning
for the entire Kanayannur near-shore zone at a total cost under \INR{}20\,000---a
fraction of the value of even a single fishing boat.
The economic argument for deployment is therefore compelling:\ the system's preventive
value far exceeds its cost within a single monsoon season if even one major event is
anticipated.

% ─────────────────────────────────────────────────────────────────────────────
\section{Supporting Disaster Preparedness Efforts}

WaveGuard~V4 integrates naturally into Kerala's existing coastal disaster management
framework.
The alert log stored in the SQLite database provides a verified, timestamped record
of sea-state events that can be submitted to the Kerala State Disaster Management
Authority (KSDMA) and INCOIS for retrospective analysis and model validation.
The system is compatible with future integration into the INCOIS Swell Surge
Forecasting Service (Section~\ref{chap:future}), enabling bidirectional information
exchange between community sensors and national agencies.

% ─────────────────────────────────────────────────────────────────────────────
\section{Academic Relevance}

WaveGuard~V4 addresses a genuinely open real-world problem documented in peer-reviewed
literature~\cite{incois2024kallakkadal, remya2016teleconnection, nair2018coastal}.
The project demonstrates practical application of:
\begin{itemize}
  \item Embedded systems design (ESP32, MEMS, GSM).
  \item IoT communication protocols (WiFi REST, GSM AT commands).
  \item Backend engineering (FastAPI, SQLite, SQLAlchemy, Pydantic).
  \item Frontend development (React.js, Vite, Recharts, Leaflet).
  \item Signal processing (RMS threshold classification).
\end{itemize}
It serves as a replicable template for future student projects involving IoT-based
environmental monitoring, and the open-source codebase and documented hardware design
facilitate direct continuation by future cohorts.

% ─────────────────────────────────────────────────────────────────────────────
\section{Mapping to Sustainable Development Goals}

WaveGuard~V4 contributes directly to three United Nations SDGs~\cite{un2015sdgs}:

\subsection*{SDG 11 – Sustainable Cities and Communities}

\textbf{Target 11.5:} ``By 2030, significantly reduce the number of deaths and the
number of people affected and substantially decrease the direct economic losses
relative to global gross domestic product caused by disasters, including water-related
disasters.''

WaveGuard~V4 reduces disaster-related casualties and property losses in coastal
communities through affordable, scalable early-warning technology that is accessible
to local panchayat administrations and fishermen's cooperatives.

\subsection*{SDG 13 – Climate Action}

\textbf{Target 13.1:} ``Strengthen resilience and adaptive capacity to climate-related
hazards and natural disasters in all countries.''

The increasing frequency and intensity of Kallakkadal events is linked to changing
meteorological patterns in the Indian Ocean~\cite{remya2016teleconnection}.
WaveGuard~V4 builds community resilience to these climate-driven hazards through
continuous near-shore monitoring that adapts to changing sea conditions without
requiring model recalibration.

\subsection*{SDG 14 – Life Below Water}

\textbf{Target 14.2:} ``By 2020, sustainably manage and protect marine and coastal
ecosystems to avoid significant adverse impacts.''

By providing early warning that prevents human encroachment into hazardous sea
conditions during swell events, WaveGuard~V4 indirectly reduces emergency rescue
operations that disturb near-shore marine ecosystems.
The system's low-power, non-invasive design (passive buoy, no acoustic or chemical
emissions) ensures minimal ecological footprint during deployment.

\begin{table}[H]
\centering
\caption{SDG Contributions of WaveGuard~V4}
\label{tab:sdg}
\begin{tabularx}{\textwidth}{cXX}
\toprule
\textbf{SDG} & \textbf{Goal} & \textbf{WaveGuard~V4 Contribution} \\
\midrule
SDG 11 & Sustainable Cities and Communities &
  Reduces disaster deaths and economic losses in coastal communities \\
\addlinespace
SDG 13 & Climate Action &
  Builds adaptive capacity to climate-driven swell surge hazards \\
\addlinespace
SDG 14 & Life Below Water &
  Non-invasive early-warning reduces human risk in near-shore marine zones \\
\bottomrule
\end{tabularx}
\end{table}
