%% chapter7.tex – Results and Observations
\chapter{Results and Observations}
\label{chap:results}

\section{System Output Examples}

\subsection{Firmware Serial Monitor Output}

A representative excerpt of the serial debug output at 115\,200~baud during a
near-shore test session:

\begin{lstlisting}[language={},caption={Serial monitor output during offshore test},label={lst:serial}]
[BOOT] WaveGuard V4 v1.0 starting...
[WIFI] Connecting to WaveGuard_AP ... OK (5.2 s)
[GPS]  Waiting for fix ... FIXED (18 s) 10.0501 N, 76.1803 E
[MPU]  Calibration loaded: off=[0.003, -0.002, -0.981]
[LOOP] RMS=0.018g  STATE=CALM      POST=OK 143ms
[LOOP] RMS=0.024g  STATE=CALM      POST=OK 141ms
[LOOP] RMS=0.047g  STATE=MODERATE  POST=OK 155ms
[LOOP] RMS=0.081g  STATE=HIGH_WAVE POST=OK 160ms
[SMS]  Sending alert to +91XXXXXXXXXX ...
[SMS]  Delivered in 9.4 s
[LOOP] RMS=0.073g  STATE=MODERATE  POST=OK 148ms
\end{lstlisting}

\subsection{FastAPI Backend JSON Response}

A sample JSON payload from \texttt{GET /readings/latest}:

\begin{lstlisting}[language={},caption={Sample API response},label={lst:api_response}]
{
  "id": 1042,
  "timestamp": "2026-04-03T11:24:07.412Z",
  "ax": -0.021, "ay": 0.007, "az": 0.978,
  "rms": 0.081,
  "state": "HIGH_WAVE",
  "lat": 10.0501,
  "lon": 76.1803,
  "battery_v": 3.81
}
\end{lstlisting}

\subsection{SMS Alert Content}

The SMS message dispatched to registered recipients on a HIGH\_WAVE event:

\begin{quote}
\ttfamily
WAVEGUARD ALERT!\\
State: HIGH WAVE\\
RMS: 0.081 g\\
Location: 10.0501N, 76.1803E\\
Time: 2026-04-03 11:24:07 UTC\\
Take precautions immediately.
\end{quote}

\section{Dashboard Observations}

The React dashboard displayed:
\begin{itemize}
  \item Live RMS gauge updating every $\approx$1~s with correct colour
        (green/yellow/red corresponding to calm/moderate/high-wave).
  \item Recharts time-series graph correctly reflecting the state transitions
        observed in the serial monitor.
  \item Leaflet map pin tracking the buoy's GPS position (stationary during
        anchored test; within 4~m of manually measured anchor coordinates).
  \item Alert log showing all three HIGH\_WAVE triggers during the pool vigorous-wave
        test, with correct timestamps.
\end{itemize}

\section{Battery Performance}

\begin{table}[H]
\centering
\caption{Battery Discharge Profile (2200 mAh, continuous operation)}
\label{tab:battery_discharge}
\begin{tabular}{rll}
\toprule
\textbf{Elapsed (h)} & \textbf{Cell Voltage (V)} & \textbf{State} \\
\midrule
0.0  & 4.15 & Full \\
1.0  & 4.01 & Normal \\
2.0  & 3.90 & Normal \\
3.0  & 3.81 & Normal \\
4.0  & 3.73 & Normal \\
5.0  & 3.65 & Normal \\
6.0  & 3.58 & Near cutoff \\
6.4  & 3.50 & MT3608 dropout; system reset \\
\bottomrule
\end{tabular}
\end{table}

The system operated normally until the MT3608 input voltage fell below its 2.0~V
minimum at $t \approx 6.4$~h, at which point the boost output dropped and the ESP32
reset.
No data corruption was observed; SQLite remained consistent after controlled shutdown.

\section{Motion Detection Accuracy}

In pool testing with a USB-ADXL345 reference accelerometer mounted at the same point:

\begin{itemize}
  \item Mean absolute error in $a_{\rm rms}$: \textbf{0.006~g}.
  \item All 50 1-second windows classified identically by WaveGuard~V4 and the
        reference system -- 100\% classification agreement on pool data.
\end{itemize}

\section{WiFi and GSM Performance Summary}

\begin{itemize}
  \item \textbf{WiFi uptime:} 99.2\% over 6.4~hour pool test (3 brief disconnects of
        $<$5~s each, auto-reconnected).
  \item \textbf{SMS delivery rate:} 96\% (48/50 triggered events delivered).
  \item \textbf{Average SMS latency:} 9.8~s (range: 8.3--11.6~s).
  \item \textbf{POST success rate:} 99.7\% (1 timeout in 320 POSTs during WiFi
        reconnect window).
\end{itemize}

\section{Observations on System Stability}

\begin{enumerate}
  \item The 1\,000~µF capacitor on the SIM800L rail eliminated all brownout resets
        during SMS transmission; prior to this fix, $\sim$30\% of SMS attempts caused
        an ESP32 brown-out reset.
  \item The 30-second SMS cooldown timer successfully prevented SMS flooding during
        sustained HIGH\_WAVE periods (e.g., vigorous pool test lasted 2~minutes with
        only 4 SMS events rather than 120).
  \item The SQLite database grew at $\approx$400~bytes per row; 6.4~hours of data
        ($\approx$23\,000 rows) occupied $\approx$9~MB, well within the backend
        server's storage capacity.
  \item GPS position jitter was $<$3~m, within the NEO-6M's rated 2.5~m CEP,
        confirming correct NMEA parsing.
\end{enumerate}
