% ============================================================
%  Chapter 3 – Background
% ============================================================
\chapter{Background}
\label{chap:background}

\section{Introduction}

This chapter provides the theoretical and technical background for the tools, protocols, and algorithms employed in WaveGuard V4.
Understanding these foundations is essential to appreciating the design decisions described in Chapter~\ref{chap:design}.

\section{FastAPI – Python Asynchronous Web Framework}

FastAPI \cite{fastapi2023} is a modern, high-performance Python web framework built on Starlette (the ASGI toolkit) and Pydantic (data validation).
Its key characteristics relevant to WaveGuard V4 are:

\begin{itemize}
  \item \textbf{ASGI concurrency:} FastAPI uses Python's \texttt{asyncio} event loop, enabling thousands of concurrent connections without thread-per-request overhead. This is critical for handling simultaneous buoy telemetry ingestion and frontend polling.
  \item \textbf{Automatic schema generation:} All API endpoints are automatically documented via OpenAPI 3.x (Swagger UI at \texttt{/docs}), reducing integration friction for hardware developers.
  \item \textbf{Pydantic validation:} Incoming request payloads are rigorously validated against typed schemas, rejecting malformed sensor data before it reaches business logic.
  \item \textbf{Dependency injection:} The \texttt{Depends()} system enables clean injection of database sessions, authentication guards, and rate limiters, avoiding global state.
\end{itemize}

\subsection{Rate Limiting and Security}

The login endpoint in WaveGuard V4 implements an in-process sliding-window rate limiter that permits at most five authentication attempts per IP address per 60-second window, mitigating brute-force attacks.
JWT-equivalent session tokens with an 8-hour TTL are issued on successful authentication and stored in a server-side dictionary, enabling instant revocation on logout.

\section{React and Vite}

\subsection{React}
React \cite{react2023} is a declarative JavaScript library for building component-based user interfaces.
Its virtual DOM reconciliation algorithm computes the minimal set of DOM mutations required for each state change, providing smooth real-time updates at 60 frames per second even with frequently changing telemetry data.

\subsection{Vite}
Vite \cite{vite2023} is a next-generation build tool that uses native ES module imports during development (Hot Module Replacement) and Rollup for production bundling.
The \texttt{vite.config.js} proxy configuration routes all \texttt{/api} requests from the React development server to the FastAPI backend (\texttt{localhost:8005}), enabling seamless cross-origin development without modifying backend CORS settings.

\subsection{Data Visualisation}
WaveGuard V4's frontend uses Recharts for rendering real-time time-series wave height and g-force plots.
The dashboard auto-refreshes sensor data every five seconds via \texttt{setInterval}-driven \texttt{fetch} calls, and alert states are colour-coded with CSS glassmorphism effects for visual clarity.

\section{ESP32 Microcontroller}

The ESP32-WROOM-32 module (Espressif Systems) \cite{espressif2023} is the central processing unit of the WaveGuard marine buoy.
Its specifications relevant to this application are summarised in Table~\ref{tab:esp32_specs}.

\begin{table}[H]
\centering
\caption{ESP32-WROOM-32 Key Specifications}
\label{tab:esp32_specs}
\begin{tabular}{ll}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Core architecture   & Xtensa LX6 dual-core, 240 MHz \\
Flash memory        & 4 MB \\
SRAM                & 520 KB \\
WiFi standard       & IEEE 802.11 b/g/n (2.4 GHz) \\
Bluetooth           & BT 4.2 / BLE \\
I2C interfaces      & 2 \\
Operating voltage   & 3.3 V \\
Deep sleep current  & 10 \(\mu\)A \\
\bottomrule
\end{tabular}
\end{table}

The ESP32 reads IMU data over I2C, computes a rolling average of RMS acceleration, classifies the sea state locally, and transmits the result via a simple HTTP GET request to the FastAPI backend.
This \textit{edge-compute} paradigm reduces network bandwidth requirements and ensures the buoy can make rudimentary local decisions even if connectivity is temporarily lost.

\section{MPU-6050 Inertial Measurement Unit}

The InvenSense MPU-6050 \cite{invensense2013} is a six-axis IMU combining a 3-axis MEMS gyroscope and a 3-axis MEMS accelerometer on a single die.
Table~\ref{tab:mpu6050_specs} summarises its key specifications.

\begin{table}[H]
\centering
\caption{MPU-6050 Key Specifications}
\label{tab:mpu6050_specs}
\begin{tabular}{ll}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Accelerometer full-scale range & \(\pm\)2g, \(\pm\)4g, \(\pm\)8g, \(\pm\)16g \\
Accelerometer sensitivity (±2g) & 16,384 LSB/g \\
Gyroscope full-scale range & \(\pm\)250, \(\pm\)500, \(\pm\)1000, \(\pm\)2000 °/s \\
Communication interface & I2C (up to 400 kHz) \\
Operating voltage & 2.375–3.46 V \\
Sample rate & Up to 1000 Hz (accel), 8000 Hz (gyro) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Wave Motion Estimation}

The buoy firmware computes the RMS of the acceleration magnitude vector over a sampling window \(N\):

\begin{equation}
  \bar{a} = \sqrt{\frac{1}{N}\sum_{i=1}^{N}\left(a_{x,i}^{2} + a_{y,i}^{2} + a_{z,i}^{2}\right)}
  \label{eq:rms_accel}
\end{equation}

where \(a_{x,i}\), \(a_{y,i}\), \(a_{z,i}\) are the three-axis accelerometer samples in units of \textit{g}.
Gravity is removed by subtracting the static bias, effectively isolating dynamic wave-induced acceleration.
The resulting value \(\bar{a}\) is transmitted to the backend as \texttt{avg\_motion}.

\section{SQLite and SQLAlchemy}

SQLite is a serverless, file-based relational database engine embedded directly into the FastAPI process.
SQLAlchemy \cite{sqlalchemy2023} provides the ORM layer, enabling Python objects to be persisted and queried without writing raw SQL.
The \texttt{BuoyReading} model stores each received telemetry packet with a UTC timestamp, enabling the historical data endpoint (\texttt{/api/history}) to serve time-series plots to the frontend.

\section{Open-Meteo API}

Open-Meteo \cite{openmeteo2022} is a free, open-source weather API providing sub-hourly forecasts derived from ECMWF's operational NWP model ensemble.
WaveGuard V4 queries the marine endpoint:

\begin{verbatim}
GET https://marine-api.open-meteo.com/v1/marine
    ?latitude=9.97&longitude=76.28
    &hourly=wave_height,wave_period,wave_direction
    &forecast_days=1
\end{verbatim}

The response JSON contains significant wave height (SWH), dominant wave period, and wave direction, updated hourly.
The backend caches this response for 4 seconds to avoid rate limiting during rapid frontend polling.

\section{Alert Fusion Algorithm}

The core intelligence of WaveGuard V4 is a simple but robust rule-based fusion function implemented in the FastAPI backend.
Given the buoy average motion \(\bar{a}\) and the satellite SWH \(H_s\), the alert level is determined as follows:

\begin{equation}
  \text{Alert} =
  \begin{cases}
    \texttt{WARNING} & \text{if } \bar{a} \geq 0.10\,g \text{ and } H_s \geq 2.5\,\text{m} \\
    \texttt{WATCH}   & \text{if } \bar{a} \geq 0.05\,g \text{ and } H_s \geq 1.5\,\text{m} \\
    \texttt{NORMAL}  & \text{otherwise}
  \end{cases}
  \label{eq:fusion}
\end{equation}

Requiring both conditions to be simultaneously satisfied eliminates single-sensor false positives: a large vessel wake may trigger the accelerometer threshold but will not elevate satellite SWH; conversely, a swell approaching from open water will be detected by both sensors.

\section{Summary}

This chapter has established the theoretical and technical foundations for WaveGuard V4.
The chosen technologies — FastAPI, React, ESP32, MPU-6050, SQLite, and Open-Meteo — collectively provide a low-cost, high-reliability, real-time sensing and alerting platform suited to near-shore coastal deployment.
