%% chapter1.tex – Introduction
%% Group No. 2 | Dept. of CSE (Data Science), MACE Kothamangalam
%% Team: Muhammed Ajmal P (MAC23CD040), Muhammed Adil PP (MAC23CD039),
%%       Mohammed Shadeed P (MAC23CD037), Sinan Rahman (LMAC23CD066)
%% Guide: Prof. Eldo P Elias | May 2026


\chapter{Introduction}

\section{Background}

Kerala's 590\,km coastline is home to more than 1.6~million fisherfolk who venture into
the sea daily for their livelihood.
The state experiences a phenomenon locally termed \textit{Kallakkadal} (roughly
``turbulent sea''), a class of destructive long-period swell that propagates from
distant storms in the Arabian Sea or the Indian Ocean and arrives at the shore without
any associated local storm activity~\cite{kumar2020kallakkadal}.
Because no visual cues precede the event, unsuspecting fishing boats and coastal
residents are caught off guard, leading to capsizings, destruction of boats, and
loss of life every monsoon season.

Conventional forecasting methods---buoys operated by the National Institute of Ocean
Technology (NIOT) and satellite-derived wave models---operate at coarse spatial
and temporal resolution, often failing to capture the hyper-local amplification that
occurs in shallow near-shore zones of 5--10~m depth.
Additionally, broadcast alerts via radio or television reach coastal communities with
insufficient lead time for the most vulnerable zones.

The rapid growth of affordable embedded computing (ESP32, Arduino), MEMS sensors
(MPU6050), and cellular connectivity (SIM800L/GPRS) makes it feasible to deploy
low-cost autonomous sensor nodes directly in near-shore waters.
Combining real-time accelerometry, onboard threshold logic, and dual-path alert
delivery (WiFi + GSM SMS) can bridge the gap between satellite-scale forecasting and
individual community-level warning.

\section{Problem Statement}

Despite advances in national-level ocean monitoring, there is a critical absence of
affordable, deployable near-shore early-warning systems for coastal Kerala.
Specific gaps include:

\begin{enumerate}[label=\roman*.]
  \item Lack of real-time, sub-kilometre wave monitoring at shallow-water deployment
        depths (5--10~m).
  \item Absence of automatic, multi-channel alert delivery (both internet-connected
        and offline GSM paths) suited to areas with intermittent connectivity.
  \item High cost and complexity of commercial buoy systems that render them
        inaccessible to local panchayat administrations and fishermen's cooperatives.
  \item No open-source, reproducible firmware and dashboard solution tailored to the
        \textit{Kallakkadal} threat profile of the Kerala coast.
\end{enumerate}

\noindent WaveGuard~V4 directly addresses all four gaps.

\section{Objectives}

The primary objectives of the WaveGuard~V4 project are:

\begin{enumerate}
  \item \textbf{Design} a waterproof IoT buoy that samples tri-axial acceleration at
        50~Hz using an MPU6050 sensor mounted on an ESP32 DevKit~v1 platform.
  \item \textbf{Implement} an onboard classification algorithm that categorises sea
        conditions as \textsc{calm}, \textsc{moderate}, or \textsc{high-wave} based
        on calibrated RMS acceleration thresholds.
  \item \textbf{Deliver} real-time alerts simultaneously via a WiFi-connected web
        dashboard and via GSM SMS through a SIM800L module.
  \item \textbf{Log} all sensor readings with GPS timestamps to a SQLite database via
        a FastAPI backend for post-event analysis.
  \item \textbf{Achieve} a total system cost under \INR{}3\,500 per unit, making
        community-scale deployment economically viable.
  \item \textbf{Sustain} continuous operation for 6--8~hours on a single 2\,200~mAh
        Li-ion cell without solar supplementation.
\end{enumerate}

\section{Scope}

The present work is scoped as follows:

\begin{itemize}
  \item \textbf{Deployment zone:} Kanayannur taluk, Ernakulam district, Kerala Coast
        (approx.\ 10.05\textdegree{}N, 76.18\textdegree{}E), 1~km offshore at
        5--10~m water depth.
  \item \textbf{Single-node prototype:} The report documents one physical buoy unit;
        multi-node mesh networking is deferred to future work.
  \item \textbf{Wave range:} Typical monsoon swell of 0.5--2~m; extreme events
        exceeding instrument range are logged as high-wave alerts.
  \item \textbf{Alert recipients:} Up to 5 pre-configured SMS recipients and any
        browser-capable device on the local WiFi network.
  \item \textbf{Software stack:} Arduino~1.8.19 firmware, FastAPI~0.115 backend,
        React~18 + Vite frontend, SQLite database.
\end{itemize}

\section{Mapping to Sustainable Development Goals}

WaveGuard~V4 contributes to three United Nations SDGs~\cite{un2015sdgs}:

\begin{description}
  \item[SDG~11 – Sustainable Cities and Communities]
        The system provides coastal communities with timely early warnings, directly
        reducing disaster risk and enhancing community resilience (Target~11.5).

  \item[SDG~13 – Climate Action]
        By monitoring climate-exacerbated coastal hazards (\textit{Kallakkadal}
        intensification under global warming), WaveGuard~V4 supports climate
        adaptation at the local level (Target~13.1).

  \item[SDG~14 – Life Below Water]
        Real-time wave intelligence helps fisherfolk avoid hazardous conditions,
        protecting both human lives and the fishing economy that depends on
        healthy marine ecosystems (Target~14.b).
\end{description}

\section{Organisation of the Report}

The remainder of this report is organised as follows.
Chapter~\ref{chap:litsurvey} reviews existing coastal monitoring systems, IoT sensor
platforms, and early-warning architectures.
Chapter~\ref{chap:background} provides technical background on the core technologies
employed.
Chapter~\ref{chap:requirements} enumerates functional, non-functional, hardware, and
software requirements.
Chapter~\ref{chap:design} describes the system design, pin configuration, power
architecture, and software implementation.
Chapter~\ref{chap:evaluation} presents the testing methodology and performance metrics.
Chapter~\ref{chap:results} documents experimental results and observations.
Chapter~\ref{chap:future} outlines directions for future enhancement.
Chapter~\ref{chap:conclusion} concludes the report.
Appendix~\ref{app:firmware} contains the firmware source code, bill of materials, and
wiring diagram.
