%% chapter2.tex – Literature Survey
%% Group No. 2 | Dept. of CSE (Data Science), MACE Kothamangalam
%% Team: Muhammed Ajmal P (MAC23CD040), Muhammed Adil PP (MAC23CD039),
%%       Mohammed Shadeed P (MAC23CD037), Sinan Rahman (LMAC23CD066)
%% Guide: Prof. Eldo P Elias | May 2026
\chapter{Literature Survey}
\label{chap:litsurvey}

This chapter reviews ten key research works related to the WaveGuard~V4 project,
covering swell surge studies, IoT-based coastal monitoring, wave forecasting, and
real-time alerting systems.
For each paper the primary advantages and limitations relevant to the present work are
identified, and a consolidated research gap table is presented at the end of the chapter.

% ─────────────────────────────────────────────────────────────────────────────
\section{Paper 1: Swell Surge (Kallakkadal) Events Along the Indian Coast}

\textbf{Authors:} ESSO--INCOIS Scientists\\
\textbf{Year:} 2024\\
\textbf{Reference:} \cite{incois2024kallakkadal}

This technical report from the Earth System Sciences Organisation -- Indian National
Centre for Ocean Information Services (ESSO--INCOIS) systematically catalogues swell
surge (locally termed \textit{Kallakkadal}) events along the Indian coastline.
The study employs ocean buoy measurements and satellite altimeter data to characterise
long-period swells (wave period $>$14~s) originating from distant storms in the Southern
Ocean and Arabian Sea.
It establishes the seasonal climatology of Kallakkadal events and quantifies their
impact on Kerala's west coast.

\begin{table}[H]
\centering
\caption{Evaluation of \cite{incois2024kallakkadal}}
\label{tab:paper1}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Aspect} & \textbf{Detail} \\
\midrule
Advantages &
  Studies swell surge events comprehensively;
  identifies long-period swells using multi-source data;
  uses both buoy observations and satellite altimetry for validation. \\
\addlinespace
Disadvantages &
  Alert dissemination remains centralised through the INCOIS portal and does
  not support localised, near-shore real-time notification;
  no community-level hardware deployment is described. \\
\bottomrule
\end{tabularx}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Paper 2: Teleconnection Between North Indian Ocean High Swell Events and
         Meteorological Conditions over the Southern Indian Ocean}

\textbf{Authors:} Remya P.\,G., Kumar R., Nair T.\,M.\,B.\\
\textbf{Year:} 2016\\
\textbf{Reference:} \cite{remya2016teleconnection}

This study establishes the teleconnective link between synoptic-scale meteorological
conditions over the Southern Indian Ocean and the occurrence of high-swell events
along the North Indian Ocean coast.
Using wave model outputs and reanalysis data, the authors show that extra-tropical
cyclones near Antarctica generate swells that propagate northward over thousands of
kilometres before impacting the Indian coast, often without any concurrent local storm
activity.

\begin{table}[H]
\centering
\caption{Evaluation of \cite{remya2016teleconnection}}
\label{tab:paper2}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Aspect} & \textbf{Detail} \\
\midrule
Advantages &
  Explains the remote-storm influence scientifically;
  establishes the oceanographic basis for Kallakkadal;
  validates the long propagation path of swell energy. \\
\addlinespace
Disadvantages &
  Purely analytical and model-based;
  no real-time detection mechanism or hardware system is proposed;
  lead time estimates are not operationalised for community warning. \\
\bottomrule
\end{tabularx}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Paper 3: Coastal Flooding Events Due to Remote Forcing Along the Southwest
         Coast of India}

\textbf{Authors:} Sheela Nair L., Remya P.\,G., et al.\\
\textbf{Year:} 2018\\
\textbf{Reference:} \cite{nair2018coastal}

This work documents specific coastal flooding events on the southwest coast of India
triggered by remotely generated long-period swells rather than local meteorological
disturbances.
Field observations combined with numerical wave modelling demonstrate that swell
dominance over wind-sea in certain seasons makes these coasts uniquely vulnerable to
flooding without obvious warning signs.

\begin{table}[H]
\centering
\caption{Evaluation of \cite{nair2018coastal}}
\label{tab:paper3}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Aspect} & \textbf{Detail} \\
\midrule
Advantages &
  Demonstrates flooding without local storms;
  emphasises swell dominance as the primary hazard mechanism;
  provides case-study evidence directly relevant to the Kerala coast. \\
\addlinespace
Disadvantages &
  Event-based post-hoc analysis only;
  no real-time sensing or alerting system is developed;
  conclusions cannot be directly translated to operational warning. \\
\bottomrule
\end{tabularx}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Paper 4: Multidirectional Swell Dynamics in the Indian Ocean}

\textbf{Authors:} Swathy Krishna P.\,S., Varghese L., et al.\\
\textbf{Year:} 2024\\
\textbf{Reference:} \cite{swathy2024multidirectional}

This paper analyses the directional characteristics and energy propagation of swell
systems across the Indian Ocean using spectral wave model outputs.
It demonstrates that swell energy arrives from multiple source regions simultaneously,
creating complex interference patterns in near-shore zones that amplify wave heights
beyond individual source predictions.

\begin{table}[H]
\centering
\caption{Evaluation of \cite{swathy2024multidirectional}}
\label{tab:paper4}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Aspect} & \textbf{Detail} \\
\midrule
Advantages &
  Studies swell direction and energy propagation quantitatively;
  useful for understanding wave behaviour in near-shore amplification zones. \\
\addlinespace
Disadvantages &
  Entirely model-based;
  no hardware implementation or near-shore buoy validation is presented;
  real-time operational use is not addressed. \\
\bottomrule
\end{tabularx}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Paper 5: Wave-Induced Coastal Flooding Along the Southwest Coast of India}

\textbf{Authors:} Ramesh Madipally, Nair L.\,S.\\
\textbf{Year:} 2023\\
\textbf{Reference:} \cite{madipally2023wave}

This study examines wave-induced coastal flooding along the southwest coast of India
using a combination of in-situ observations and numerical simulations.
The authors quantify the contribution of wave run-up and overtopping to coastal
inundation and identify near-shore bathymetric features that exacerbate flooding.

\begin{table}[H]
\centering
\caption{Evaluation of \cite{madipally2023wave}}
\label{tab:paper5}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Aspect} & \textbf{Detail} \\
\midrule
Advantages &
  Provides quantitative wave run-up estimates for the southwest Indian coast;
  identifies bathymetric amplification zones relevant to near-shore warning
  placement. \\
\addlinespace
Disadvantages &
  Limited to simulation and post-event analysis;
  computational requirements restrict real-time use;
  no wearable or embedded IoT sensing component. \\
\bottomrule
\end{tabularx}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Paper 6: Wave Monitoring Along Nearshore Buoy Network}

\textbf{Authors:} Balakrishnan Nair T.\,M., et al.\\
\textbf{Year:} 2025\\
\textbf{Reference:} \cite{balakrishnannair2025buoy}

This recent paper describes the design and operation of a nearshore buoy network for
continuous wave monitoring along the Indian coastline.
The network employs directional wave rider buoys connected via GPRS to a central
data server, providing near-real-time wave height, period, and direction data.
The system improves forecast accuracy by assimilating measured data into numerical
wave models.

\begin{table}[H]
\centering
\caption{Evaluation of \cite{balakrishnannair2025buoy}}
\label{tab:paper6}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Aspect} & \textbf{Detail} \\
\midrule
Advantages &
  Demonstrates real-time buoy-based wave monitoring operationally;
  improves forecast accuracy through data assimilation;
  directly validates the concept of in-situ near-shore sensing. \\
\addlinespace
Disadvantages &
  High installation and ongoing maintenance cost makes community-scale
  deployment financially infeasible;
  no low-cost embedded solution is proposed. \\
\bottomrule
\end{tabularx}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Paper 7: Implementation of Altimeter Data Assimilation on Regional Wave
         Forecasting Systems}

\textbf{Authors:} Sirisha P., Remya P.\,G., Srinivas K.\\
\textbf{Year:} 2023\\
\textbf{Reference:} \cite{sirisha2023altimeter}

This paper presents the implementation of satellite altimeter data assimilation into a
regional wave forecasting model (WAVEWATCH III) for improved significant wave height
prediction.
Assimilation of Jason-3 altimeter tracks into the model reduces forecast error by up
to 30\% for the peak swell season.

\begin{table}[H]
\centering
\caption{Evaluation of \cite{sirisha2023altimeter}}
\label{tab:paper7}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Aspect} & \textbf{Detail} \\
\midrule
Advantages &
  Improves wave forecast accuracy using satellite data;
  directly useful for large-scale swell surge prediction;
  validates satellite altimetry as a reliable data source. \\
\addlinespace
Disadvantages &
  Requires complex high-performance computational infrastructure;
  not suitable for embedded or low-cost community deployment;
  latency from satellite overpass to corrected forecast is too long for
  near-shore real-time alerting. \\
\bottomrule
\end{tabularx}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Paper 8: Numerical Simulation of Coastal Waves Using SWAN and MIKE Models}

\textbf{Authors:} Pakhirehzan M., et al.\\
\textbf{Year:} 2024\\
\textbf{Reference:} \cite{pakhirehzan2024numerical}

This paper applies two widely used coastal wave simulation tools---SWAN (Simulating
WAves Nearshore) and MIKE 21 SW---to model wave transformation in shallow coastal
areas.
Comparison of model outputs with observed buoy data shows good agreement for
significant wave height and mean wave period under various swell and sea-state
conditions.

\begin{table}[H]
\centering
\caption{Evaluation of \cite{pakhirehzan2024numerical}}
\label{tab:paper8}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Aspect} & \textbf{Detail} \\
\midrule
Advantages &
  Uses validated numerical models widely adopted in coastal engineering practice;
  demonstrates near-shore wave transformation processes that justify the
  shallow-water buoy placement strategy in WaveGuard~V4. \\
\addlinespace
Disadvantages &
  Not suitable for real-time alerting due to computational overhead;
  requires detailed bathymetric and meteorological input data;
  model setup and calibration demand specialised expertise. \\
\bottomrule
\end{tabularx}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Paper 9: Data-Driven Models for Significant Wave Height Forecasting}

\textbf{Authors:} Durap A., et al.\\
\textbf{Year:} 2024\\
\textbf{Reference:} \cite{durap2024waveheight}

This study develops and evaluates machine-learning-based models---including LSTM neural
networks and XGBoost regressors---for forecasting significant wave heights using
historical wave buoy and meteorological data.
The models demonstrate competitive performance against physics-based numerical models
with dramatically reduced computation time after training.

\begin{table}[H]
\centering
\caption{Evaluation of \cite{durap2024waveheight}}
\label{tab:paper9}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Aspect} & \textbf{Detail} \\
\midrule
Advantages &
  Uses machine learning for wave height prediction;
  shows the potential of intelligent, data-driven forecasting to complement
  physics-based models;
  motivates future ML integration in WaveGuard~V4. \\
\addlinespace
Disadvantages &
  Requires large historical datasets for training;
  model generalisation to unseen coastal geometries is limited;
  no embedded hardware or real-time alert delivery is described. \\
\bottomrule
\end{tabularx}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Paper 10: Coastal Flood Inundation Forecasting and Risk Assessment}

\textbf{Authors:} Raam Balaji V., et al.\\
\textbf{Year:} 2024\\
\textbf{Reference:} \cite{balaji2024inundation}

This paper presents an integrated framework for forecasting coastal flood inundation
extent and assessing flood risk for populated coastal areas.
It combines wave run-up models, coastal DEM data, and probabilistic risk metrics to
produce inundation maps that inform disaster preparedness planning.

\begin{table}[H]
\centering
\caption{Evaluation of \cite{balaji2024inundation}}
\label{tab:paper10}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Aspect} & \textbf{Detail} \\
\midrule
Advantages &
  Forecasts coastal inundation extent with spatial detail;
  useful for disaster preparedness and evacuation planning;
  demonstrates value of integrated coastal risk tools. \\
\addlinespace
Disadvantages &
  Not localised to individual community-level zones;
  no real-time hardware sensing component;
  primarily a planning tool rather than an operational warning system. \\
\bottomrule
\end{tabularx}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Research Gap and WaveGuard Contribution}

Table~\ref{tab:litgap} synthesises the key capabilities reviewed across all ten papers
and identifies the gap that WaveGuard~V4 fills.

\begin{table}[H]
\centering
\caption{Comparison of reviewed works with WaveGuard~V4}
\label{tab:litgap}
\begin{tabularx}{\textwidth}{Xccccc}
\toprule
\textbf{Work} &
\textbf{\shortstack{Near-shore\\sensing}} &
\textbf{\shortstack{Real-time\\alert}} &
\textbf{\shortstack{Low\\cost}} &
\textbf{\shortstack{GSM /\\SMS}} &
\textbf{\shortstack{Web\\dashboard}} \\
\midrule
INCOIS 2024~\cite{incois2024kallakkadal}
  & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{51} \\
Remya 2016~\cite{remya2016teleconnection}
  & \ding{55} & \ding{55} & n/a       & \ding{55} & \ding{55} \\
Sheela Nair 2018~\cite{nair2018coastal}
  & \ding{55} & \ding{55} & n/a       & \ding{55} & \ding{55} \\
Swathy Krishna 2024~\cite{swathy2024multidirectional}
  & \ding{55} & \ding{55} & n/a       & \ding{55} & \ding{55} \\
Madipally 2023~\cite{madipally2023wave}
  & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} \\
Balakrishnan Nair 2025~\cite{balakrishnannair2025buoy}
  & \ding{51} & \ding{51} & \ding{55} & \ding{55} & \ding{51} \\
Sirisha 2023~\cite{sirisha2023altimeter}
  & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} \\
Pakhirehzan 2024~\cite{pakhirehzan2024numerical}
  & \ding{55} & \ding{55} & n/a       & \ding{55} & \ding{55} \\
Durap 2024~\cite{durap2024waveheight}
  & \ding{55} & \ding{55} & n/a       & \ding{55} & \ding{55} \\
Raam Balaji 2024~\cite{balaji2024inundation}
  & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} \\
\midrule
\textbf{WaveGuard~V4 (this work)}
  & \textbf{\ding{51}} & \textbf{\ding{51}} & \textbf{\ding{51}}
  & \textbf{\ding{51}} & \textbf{\ding{51}} \\
\bottomrule
\end{tabularx}
\end{table}

\noindent The reviewed literature clearly establishes the scientific and operational
importance of Kallakkadal monitoring, yet no existing low-cost system simultaneously
delivers near-shore MEMS-based wave sensing, real-time threshold classification,
GSM SMS alerting, and a web dashboard at community-affordable cost.
WaveGuard~V4 directly addresses all five of these identified gaps, leveraging commodity
IoT hardware and open-source software to bridge the gap between national-scale
forecasting and localised community warning.
